I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-03 Paper/Paper_noProof.tex
llncs
Writing
cav

\documentclass[8pt]{article}
\usepackage{amsmath, amsthm, amssymb, amsfonts}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newcommand{\ReLU}{\mathrm{ReLU}}

@@

\subsection{The neural network and Verification problem}

@@

In this paper, we focus on full-connected neural network. The network has weights and bias to send the input to output along hidden layers. In formula, the input layer of a neural network is a vector $x$ in $\mathbb{R}^{d_0}$, and the $i$-th hidden layer  is a vector in $\mathbb{R}^{d_i}$, and the output layer can be a vector in $\mathbb{R}^{d'}$ or a scale. The weights, bias and activation functions decide propagate the from previous to the next layer. In formula, from layer $l_{i-1}$ to layer $l_{i}$, the weight $W^i$ is matrix of $d_i\times d_{i-1}$, the bias is a vector $b^i$ in $\mathbb{R}^{d_i}$, and the activation function is $\sigma$, then  if the $i-1$-th layer is $\hat{z}^{(i-1)}$, then the value of $i$-th layer is computed by: \begin{align*}
	{z}^{i} &= W^i\cdot \hat{z}^{(i-1)}+ b^i\\
	\hat{z}^{i}(n) &= \sigma({z}^i(n)).
\end{align*} The vector $\hat{z}$ is called pre-activation values, while $z$ is called post-activation values, and $z^{(i)}_j$ is used to call the $j$-th neuron in the $i$-th layer. In our style, we also call neurons \emph{nodes} and use $a,b,c,d$ to denote them. We use $W_{ab}$ to denote the weight from neuron $b$ to $a$. We use $x$ to denote the vector of input and  $f(x)$ to denote the output.

The verification problem is to determine whether the output of a neural network will be affected under small perturbations to the input. In formula, if we use a distance $d$, and if the $x_0$ is a ordinary input, then the considered domain $\mathcal{D}=\{x: d(x,x_0)<\epsilon\}$ where $\epsilon$ is the parameter measuring the size of perturbation. And the problem will be \begin{align*}
	\forall x\in\mathcal{D} \   f(x) = f(x_0).
\end{align*} In some cases, the output is a vector but the aim to get the label of dimension with the minimal value. In this case, the problem can be written as:\begin{align*}
\forall x \in\mathcal{D} \  \min f(x) = \min f(x_0)
\end{align*}

If so, the question of verification can turn to the following optimization question: \begin{align*}
	\min f(x) \ s.t. {z}^{i} &= W^i\cdot \hat{z}^{(i-1)}+ b^i\\
	\hat{z}^{i}(n) &= \sigma({z}^i(n)), x\in\mathcal{D}.
\end{align*}

In this paper, we only consider $\ReLU$ function as the activation function: $\sigma(a)=\ReLU(a)=\max(0,a)$. We consider $L^{\infty}$ norm, that is $d(x,x_0)=\max |x(n)-x_0(n)|$, the max value of distance of each dimension.

@@

It is known in literature that if $x=\ReLU(y)$, and $y$ has its upper and lower bounds $u>0$ and $l<0$ (we call this unstable node; and if $u\geq 0$ or $l\leq 0$ then it is called stable node, and does not need MILP formulation), then this function can be formulated in MILP with one integer variable $a$ valued in ${0,1}$ (i.e., a binary variable) by:

@@

It is a standard method to compute the lower bounds and upper bounds of neurons in a pure $\ReLU$ activation network by MILP optimization, because from previous subsection, the verification problem is equivalent to an optimization problem.

@@

\section{Compensate, Diamond and Path chosen}

In this section, we will develop the process of open node chosen.

@@

In this subsection, we will explore the key factor for the loss of accuracy in our LP/MILP models. The answer is the \emph{compensating pair}.

In above figure, $a$ is the input neuron, $bc,b'c'$ are two nodes in the hidden layer, before and after the activation function, and $d$ is the unique output neuron. The numbers next to the arrows are the weights. So, $W_{ba}=1$ and $W_{b'a}=-1$, $W_{dc}=W_{dc'}=1$. The paths, $a$ to $bc$ to $d$, and $a$ to $b'c'$ to $d$, is the so called compensating pair, or a Diamond. The key point is that, the products of all weights in the paths, have two different signs: along $bc$, the product is (strictly) positive, while along $b'c'$, the product is (strictly) negative.

@@

\subsection{Experiment setting}

@@

\subsection{Experiments Result}

The following chart is the summary of 5 experiments.

>>>>

\documentclass{llncs}
\pagestyle{plain}

%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usetikzlibrary{positioning, arrows.meta,calc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{amsmath, amsthm, amssymb, amsfonts}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{corollary}{Corollary}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}

\newcommand{\ReLU}{\mathrm{ReLU}}

@@

We follow some notions in PRIMA's and Beta-Crown's papers: we use lower case Latin for scalars, bold for vectors
$\boldsymbol{z}$, capitalized bold for matrices $\boldsymbol{W}$. In this paper, we will not use tensors with a dimension higher than matrices: those will be flattened.

\subsection{Neural Network and Verification}

@@

We focus on full-connected neural network. The network has weights and bias to send the input to output along hidden layers. In formula, the input layer of a neural network is a vector $\boldsymbol{x}$ in $\mathbb{R}^{d_0}$, and the $i$-th hidden layer  is a vector in $\mathbb{R}^{d_i}$, and the output layer is a vector in $\mathbb{R}^{d'}$ or a scale. The weights, bias and activation functions decide propagate the from previous to the next layer. In formula, from layer $l_{i-1}$ to layer $l_{i}$, the weight $\boldsymbol{W}^i$ is matrix of $d_i\times d_{i-1}$, the bias is a vector $\boldsymbol{b}^i$ in $\mathbb{R}^{d_i}$, and the activation function is $\sigma$, then  if the $i-1$-th layer is $\hat{\boldsymbol{z}}^{(i-1)}$, then the value of $i$-th layer is computed by: \begin{align*}
	{\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ \boldsymbol{b}^i\\
	\hat{\boldsymbol{z}}^{i}(n) &= \sigma({\boldsymbol{z}}^i(n)).
\end{align*} The vector $\hat{\boldsymbol{z}}$ is called post-activation values, and $\boldsymbol{z}$ is called pre-activation values, and $\boldsymbol{z}^{(i)}_j$ is used to call the $j$-th neuron in the $i$-th layer. In our style, we also call neurons \emph{nodes} and use $a,b,c,d$ to denote them. We use $W_{ab}$ to denote the weight from neuron $b$ to $a$. We use $\boldsymbol{x}$ to denote the vector of input and  $f(\boldsymbol{x})$ to denote the output.

The verification problem is to determine whether the output of a neural network will be affected under small perturbations to the input. In formula, if we use a distance $d$, and if the $\boldsymbol{x}_0$ is a ordinary input, then the considered domain $\mathcal{D}=\{x: d(\boldsymbol{x},\boldsymbol{x}_0)<\varepsilon\}$ where $\varepsilon$ is the parameter measuring the size of perturbation. And the problem will be \begin{align*}
	\forall \boldsymbol{x}\in\mathcal{D} \   f(\boldsymbol{x}) = f(\boldsymbol{x}_0).
\end{align*} In some cases, the output is a vector but the aim to get the label of dimension with the minimal value. In this case, the problem can be written as:\begin{align*}
\forall \boldsymbol{x} \in\mathcal{D} \  \min f(\boldsymbol{x}) = \min f(\boldsymbol{x}_0)
\end{align*}

If so, the question of verification can turn to the following optimization question: \begin{align*}
	\min f(\boldsymbol{x}) \ s.t. {\boldsymbol{z}}^{i} &= \boldsymbol{W}^i\cdot \hat{\boldsymbol{z}}^{(i-1)}+ b^i\\
	\hat{\boldsymbol{z}}^{i}(n) &= \sigma({\boldsymbol{z}}^i(n)), \boldsymbol{x}\in\mathcal{D}.
\end{align*}

In this paper, we only consider $\ReLU$ function as the activation function: $\sigma(a)=\ReLU(a)=\max(0,a)$. We will consider $L^{\infty}$ norm, that is $$d(\boldsymbol{x},\boldsymbol{x}_0)=\max |\boldsymbol{x}(n)-\boldsymbol{x}_0(n)|,$$ the max value of distance of each dimension.

@@

In literature, many methods have been used to approximate $\ReLU$ functions. One of them is MILP (or MIP, Mixed Integer Linear Programming).

For a activation $x=\ReLU(y)$, if we already know $y's$ rough upper and lower bounds $u,l$, then if $u>0$ and $l<0$, we call $y$ is unstable; otherwise, $y$ is stable. In practice, stable node is easier because $\ReLU$ function becomes a linear function on this node. The key problem is to deal with unstable nodes.

In literature, if $x=\ReLU(y)$, and $y$ is unstable, then this function can be formulated in MILP with one integer variable $a$ valued in ${0,1}$ (i.e., a binary variable) by:

@@

MILP becomes one of standard methods to compute the lower bounds and upper bounds of neurons in a pure $\ReLU$ activation network. Combining with the discussion from previous subsection, the verification problem is equivalent to an MILP optimization problem.

@@

\section{Compensate, Diamond and Node Chosen}

In this section, we will explore the key factor for the loss of accuracy in our LP/MILP models: \emph{compensating pair}. And we will develop a process of open node chosen by the principle of \emph{Compensating Pairs}.

@@

The basic mode of compensating pair is as follows:

\vspace*{2ex}

\begin{tikzpicture}
	\node[circle, draw= purple, thick, minimum width = 20,
	minimum height = 20] (input1) {$a$};

	% Hidden layers
	\node[circle, draw= blue, thick, minimum width = 20,
	minimum height = 20] (hidden1) at ($(input1) + (2,1)$) {$b$};
	\node[circle, draw= blue, thick] (hidden2) at ($(input1) + (2,-1)$) {$b'$};

	\node[circle, draw= blue, thick, minimum width = 20,
	minimum height = 20] (hidden3) at ($(input1) + (4,1)$){$c$};
	\node[circle, draw= blue, thick] (hidden4) at ($(input1) + (4,-1)$) {$c'$};

	% Output layer
	\node[circle, draw= blue, thick, minimum width = 20,
	minimum height = 20] (output) at ($(input1) + (6,0)$){$d$};

	% Connections
	\draw[->,thick,draw= red] (input1) -- (hidden1) node[midway, below] {$1$};
	\draw[->,thick,draw= green] (input1) -- (hidden2)node[midway, below] {$-1$};

	\draw[->,thick,draw= red] (hidden1) -- (hidden3) node[midway, below] {$\ReLU$};
	\draw[->,thick,draw= green] (hidden2) -- (hidden4) node[midway, below] {$\ReLU$};

	\draw[->,thick,draw= red] (hidden3) -- (output)node[midway, below] {$1$};
	\draw[->,thick,draw= green] (hidden4) -- (output)node[midway, below] {$1$};
\end{tikzpicture}

\vspace*{2ex}

In this figure, $a$ is the input neuron; $bc,b'c'$ are two nodes in the hidden layer, pre and post activation; and $d$ is the unique output neuron. The numbers next to the arrows are the weights. So, $W_{ba}=1$ and $W_{b'a}=-1$, $W_{dc}=W_{dc'}=1$. The paths, $a$ to $bc$ to $d$, and $a$ to $b'c'$ to $d$, is the so called \emph{Compensating Pair}. Because this figure looks like a diamond, it is also called a Diamond.

The key point is that, the products of all weights in the paths, have two different signs: along $bc$, the product is (strictly) positive, while along $b'c'$, the product is (strictly) negative.

@@

\section{Verification Framework}

@@

\subsection{Experiments about Node Chosen}

The first group of experiments is to test the direct effect of node chosen by compensating pair, comparing to random chosen. We use the average of uncertainty of all nodes in a layer to measure the effect of node chosen: for a node $a$, its uncertainty $UN(a) = UB(a)-LB(a)$, and the average uncertainty of a layer $l$ is $$UN(l) = \dfrac{\sum_{a\in l}UN(a)}{|l|}.$$

The test neural network is MNIST fully-connected $6\times 100$, and the test image is the image 59 in the standard MNIST image database. This image can have more unstable nodes in each layer.

\subsubsection*{Node chosen in second hidden layer}

The following table shows the average uncertainty of the second hidden layer when choosing different numbers of open nodes by random or by compensating pairs.

\subsubsection*{Node chosen in the third hidden layer}

The following table shows the average uncertainty of the third hidden layer when choosing different numbers of open nodes, by random or by compensating pairs.

\subsection{Experiments Results}

The following table is the summary of 5 experiments.
